pema application

### 2023-02-09 paper/manuscript.Rmd
add results files for knitting

<span id = "openingparagraph">A common application of meta-analysis is to summarize existing bodies of literature.
A crucial challenge is that there is often substantial heterogeneity between studies,
because similar research questions are studied in different labs,
sampling from different populations,
and using different study designs, instruments, and methods.
Any of those between-studies differences can introduce *systematic heterogeneity* in observed effect sizes.
Suspected causes of systematic heterogeneity can either be used as exclusion criteria,
or controlled for using *meta-regression* [see @lopez-lopezEstimationPredictivePower2014].
The latter approach provides an opportunity to learn which factors impact the effect size
found.
However, a limitation of meta-regression is that it requires a relatively high number of cases (studies) per parameter to obtain sufficient statistical power.
In applied meta-analyses, the number of available studies is often low [@rileyInterpretationRandomEffects2011].
This introduces a risk of overfitting,
which results in uninterpretable model parameters [@hastieElementsStatisticalLearning2009].
In extreme cases, the ratio of cases to parameters can be so low that the model is not (empirically) identified,
resulting in non-convergence [@akaikeNewLookStatistical1974].
Accounting for between-studies heterogeneity thus poses a non-trivial challenge to classic meta-analytic methods.
The risk of arriving at false-positive conclusions
when there are many potential moderators is so ubiquitous that it was referred to as the "primary pitfall" in meta-regression [@thompsonHowShouldMetaregression2002].
The present study introduces a novel method to overcome this pitfall by imposing Bayesian regularizing (LASSO and regularized horseshoe) priors on the regression coefficients.
These priors shrink the effect of irrelevant predictors towards zero while leaving important predictors relatively unaffected.
The result is a sparse model with fewer non-zero parameters,
which benefits model convergence, reduces overfitting, and helps identify relevant between-study differences.
</span>

@@

Meta-regression models are most commonly estimated using weighted least squares,
with weights determined according to the aforementioned random effects model,
where residual heterogeneity is estimated using restricted maximum likelihood [@pattersonRecoveryInterblockInformation1971; @viechtbauerBiasEfficiencyMetaAnalytic2005].
This approach, henceforth referred to as RMA, has low bias, which means that, across hypothetical replications of a study, the average values of model parameters are close to their true population values [@panityakulEstimatingResidualHeterogeneity2013].
Low bias comes at the cost of higher variance, however,
which means that the estimated values of population parameters vary more from one hypothetical replication to the next.
This phenomenon is known as the *bias-variance trade-off*.
In general, an estimator with low bias and high variance produces results that generalize less well to new data than an estimator with high bias and low variance.
Regularized approaches to regression intentionally increase bias in order to reduce variance.
This is a sensible approach in the context of small samples, which are common in meta-analyses,
because small samples incur a high risk of overfitting,
and typically have relatively high levels of multicollinearity,
due to the higher probability that extreme values on one moderator coincide with extreme values on another.
In such cases, regularized regression reduces the risk of overfitting and increase generalizability of the results [see @hastieElementsStatisticalLearning2009].

<span id = "reftibshirani">To understand how regularized regression introduces bias, consider a comparison between ordinary least squares regression and LASSO regression [for a more elaborate introduction, see @tibshiraniRegressionShrinkageSelection2011].</span>
Ordinary least squares regression (OLS) estimates model parameters by minimizing the Residual Sum of Squares (RSS) of the outcome variable, given by the formula below.
In this equation, $n$ is the number of participants; $y_i$ is the outcome variable, and $x$ is one of $p$ predictor variables.
\begin{align}
RSS=\sum_{i=1}^{n}(y_{i} - \beta_{0} - \sum_{j=1}^{p}\beta_{j}x_{ij})^2
\end{align}
The resulting parameter estimates perfectly describe linear relations in the present data set, but generalize less well to new data.
Regularized regression biases parameter estimates towards zero by adding a penalty term to the RSS.
Most common is the LASSO penalty, which consists of the sum of the absolute regression coefficients, or the L1 norm [@hastieElementsStatisticalLearning2009].
As the LASSO penalty is a function of the regression coefficients, it increases when they get bigger.
This incentivizes the optimizer to keep the regression coefficients as small as possible.
The amount of regularization can be controlled by multiplying the penalty by a tuning parameter, $\lambda$.
If $\lambda$ is zero, the shrinkage penalty has no impact.
As $\lambda$ increases,
all coefficients shrink towards zero, ultimately producing the null model.
Cross-validation is often used to find the optimal value for the penalty parameter $\lambda$.
The LASSO penalized residual sum of squares is given by:
\begin{align}
PRSS= RSS + \lambda \sum_{j=1}^{p}|\beta_{j}| (\#eq:prss)
\end{align}
<span id = "whylasso">Note that many other regularizing penalties exist.
This introduction focuses on the LASSO penalty because it is most ubiquitous, easy to understand,
and has an analogue in Bayesian estimation,
as explained in the next section.</span>

@@

We performed a simulation study to validate the BRMA algorithm.
As a benchmark for comparison, we used random effects meta-regression with restricted maximum likelihood estimation [RMA, @viechtbauerConductingMetaanalysesMetafor2010],
which is the current state-of-the-art in the field.
We evaluated the algorithms' predictive performance in new data, 
ability to perform variable selection,
and ability to recover population parameters.
Our research questions are whether BRMA offers a performance advantage over RMA in terms of any of these indicators,
and which prior (LASSO versus regularized horseshoe) is to be preferred.
<span id = "defaultvalues">For both Bayesian priors, we used default values proposed in prior literature, see Table \@ref(tab:tabprior).
Default values for the LASSO prior were based on @ParkCasella2008,
and default values for the regularized horseshoe prior were based on @PiironenVehtari2017a.</span>
The simulation code is available in a version-controlled repository at <https://github.com/cjvanlissa/pema>, in the directory `pema/simulatie_2021`.

@@

<span id = "metareglimits">Before conducting meta-regression,
researchers should be aware of its limitations [see @thompsonHowShouldMetaregression2002].
These can be subdivided into four categories: 1) the curse of dimensionality and its corrolary implications for multicollinearity; 2) the ecological fallacy; 3) limited information on moderator variables, including missing data and restrictions of range.
BRMA seeks to address the first of these limitations,
because the problems that arise from meta-analyzing small and heterogeneous bodies of literature are so ubiquitous that they have been referred to as the primary pitfall in meta-regression [@thompsonHowShouldMetaregression2002].
Nonetheless, all applicable limitations should be acknowledged in the resulting publication.</span>

With regard to the planning and design of a BRMA meta-analysis, consider explicitly mentioning the intended use of BRMA in a preregistered analysis protocol - either as primary analysis technique, or as a contingency in the case of model non-convergence or multicollinearity.
Note that BRMA is suitable for both confirmatory hypothesis tests and for exploratory analyses to ensure that no important effects were missed.
Both approaches can be included in a preregistration [see @vanlissaComplementingPreregisteredConfirmatory2022].
With regard to data extraction,
it is important to strike a balance between inclusiveness and selectivity when coding moderators [@thompsonHowShouldMetaregression2002].
Moderators may include theoretically relevant factors and methodological ones, such as sample demographics, methods, instruments, study quality, and publication type.
A key challenge is that moderators may not always be reported.
The best way to handle missing data is by recovering the relevant information
by contacting authors or comparing different publications on the same data.
If data remains missing, users can use multiple imputation, which is a best practice for handling missingness (see Applied example).
Finally, effect sizes and their variances must be computed using suitable methods;
many of which are available in the R package `metafor` [@viechtbauerConductingMetaanalysesMetafor2010].

@@

With regard to publication, we highly recommend making the data and code for the meta-analysis publicly available.
One way to do this is by creating a reproducible research repository,
for example, using the Workflow for Reproducible Code in Science [WORCS, @vanlissaWORCSWorkflowOpen2021].
Transparency allows readers and reviewers to verify that methods were correctly applied,
which bolsters confidence in the results.
Others can easily perform sensitivity analyses by changing the analysis code.
Sharing data allows the meta-analysis to be updated in the future,
which increases its reuse value.
Finally, sharing the model object (or code to reproduce it) allows others to obtain predictions for the expected effect size of a new study on the same topic.
These predictions can be used to conduct power analysis for future research.
To this end, researchers can simply enter their planned
design (or several alternative designs) as new lines of data, using the codebook of the
original meta-analysis, and use the published BRMA model to calculate the
predicted effect size for a study with these specifications.

>>>>

<span id = "openingparagraph">A common application of meta-analysis is to summarize existing bodies of literature.
A crucial challenge is that there is often substantial heterogeneity between studies,
because similar research questions are studied in different labs,
sampling from different populations,
and using different study designs, instruments, and methods.
Any of those between-studies differences can introduce *systematic heterogeneity* in observed effect sizes.
Suspected causes of systematic heterogeneity can either be used as exclusion criteria,
or controlled for using *meta-regression* [@lopez-lopezEstimationPredictivePower2014].
The latter approach provides an opportunity to learn which factors impact the effect size
found.
However, a limitation of meta-regression is that it requires a relatively high number of cases (studies) per parameter to obtain sufficient statistical power.
In applied meta-analyses, the number of available studies is often low [@rileyInterpretationRandomEffects2011].
This introduces a risk of overfitting,
which results in uninterpretable model parameters [@hastieElementsStatisticalLearning2009].
In extreme cases, the ratio of cases to parameters can be so low that the model is not (empirically) identified,
resulting in non-convergence [@akaikeNewLookStatistical1974].
Accounting for between-studies heterogeneity thus poses a non-trivial challenge to classic meta-analytic methods.
The risk of arriving at false-positive conclusions
when there are many potential moderators is so ubiquitous that it was referred to as the "primary pitfall" in meta-regression [@thompsonHowShouldMetaregression2002].
The present study introduces a novel method to overcome this pitfall by imposing Bayesian regularizing (LASSO and regularized horseshoe) priors on the regression coefficients.
These priors shrink the effect of irrelevant predictors towards zero while leaving important predictors relatively unaffected.
The result is a sparse model with fewer non-zero parameters,
which benefits model convergence, reduces overfitting, and helps identify relevant between-study differences.
</span>

@@

Meta-regression models are most commonly estimated using weighted least squares,
with weights determined according to the aforementioned random effects model,
where residual heterogeneity is estimated using restricted maximum likelihood [@pattersonRecoveryInterblockInformation1971; @viechtbauerBiasEfficiencyMetaAnalytic2005].
This approach, henceforth referred to as RMA, has low bias, which means that, across hypothetical replications of a study, the average values of model parameters are close to their true population values [@panityakulEstimatingResidualHeterogeneity2013].
Low bias comes at the cost of higher variance, however,
which means that the estimated values of population parameters vary more from one hypothetical replication to the next.
This phenomenon is known as the *bias-variance trade-off*.
In general, an estimator with low bias and high variance produces results that generalize less well to new data than an estimator with high bias and low variance.
Regularized approaches to regression intentionally increase bias in order to reduce variance.
This is a sensible approach in the context of small samples, which are common in meta-analyses,
because small samples incur a high risk of overfitting,
and typically have relatively high levels of multicollinearity,
due to the higher probability that extreme values on one moderator coincide with extreme values on another.
In such cases, regularized regression reduces the risk of overfitting and increase generalizability of the results [@hastieElementsStatisticalLearning2009].

<span id = "reftibshirani">To understand how regularized regression introduces bias, consider a comparison between ordinary least squares regression and LASSO regression; for a more elaborate introduction, see [@tibshiraniRegressionShrinkageSelection2011].</span>
Ordinary least squares regression (OLS) estimates model parameters by minimizing the Residual Sum of Squares (RSS) of the outcome variable, given by the formula below.
In this equation, $n$ is the number of participants; $y_i$ is the outcome variable, and $x$ is one of $p$ predictor variables.
\begin{align}
RSS=\sum_{i=1}^{n}(y_{i} - \beta_{0} - \sum_{j=1}^{p}\beta_{j}x_{ij})^2
\end{align}
The resulting parameter estimates perfectly describe linear relations in the present data set, but generalize less well to new data.
Regularized regression biases parameter estimates towards zero by adding a penalty term to the RSS.
Most common is the LASSO penalty, which consists of the sum of the absolute regression coefficients, or the L1 norm [@hastieElementsStatisticalLearning2009].
As the LASSO penalty is a function of the regression coefficients, it increases when they get bigger.
This incentivizes the optimizer to keep the regression coefficients as small as possible.
The amount of regularization can be controlled by multiplying the penalty by a tuning parameter, $\lambda$.
If $\lambda$ is zero, the shrinkage penalty has no impact.
As $\lambda$ increases,
all coefficients shrink towards zero, ultimately producing the null model.
Cross-validation is often used to find the optimal value for the penalty parameter $\lambda$.
The LASSO penalized residual sum of squares is given by:
\begin{align}
PRSS= RSS + \lambda \sum_{j=1}^{p}|\beta_{j}| (\#eq:prss)
\end{align}
<span id = "whylasso">Note that many other regularizing penalties exist.
This introduction focuses on the LASSO penalty because it is most ubiquitous, easy to understand,
and has an analogue in Bayesian estimation,
as explained in the next section.</span>

@@

We performed a simulation study to validate the BRMA algorithm.
As a benchmark for comparison, we used random effects meta-regression with restricted maximum likelihood estimation (RMA) [@viechtbauerConductingMetaanalysesMetafor2010],
which is the current state-of-the-art in the field.
We evaluated the algorithms' predictive performance in new data, 
ability to perform variable selection,
and ability to recover population parameters.
Our research questions are whether BRMA offers a performance advantage over RMA in terms of any of these indicators,
and which prior (LASSO versus regularized horseshoe) is to be preferred.
<span id = "defaultvalues">For both Bayesian priors, we used default values proposed in prior literature, see Table \@ref(tab:tabprior).
Default values for the LASSO prior were based on @ParkCasella2008,
and default values for the regularized horseshoe prior were based on @PiironenVehtari2017a.</span>
The simulation code is available in a version-controlled repository at <https://github.com/cjvanlissa/pema>, in the directory `pema/simulatie_2021`.

@@

<span id = "metareglimits">Before conducting meta-regression,
researchers should be aware of its limitations [@thompsonHowShouldMetaregression2002].
These can be subdivided into four categories: 1) the curse of dimensionality and its corrolary implications for multicollinearity; 2) the ecological fallacy; 3) limited information on moderator variables, including missing data and restrictions of range.
BRMA seeks to address the first of these limitations,
because the problems that arise from meta-analyzing small and heterogeneous bodies of literature are so ubiquitous that they have been referred to as the primary pitfall in meta-regression [@thompsonHowShouldMetaregression2002].
Nonetheless, all applicable limitations should be acknowledged in the resulting publication.</span>

With regard to the planning and design of a BRMA meta-analysis, consider explicitly mentioning the intended use of BRMA in a preregistered analysis protocol - either as primary analysis technique, or as a contingency in the case of model non-convergence or multicollinearity.
Note that BRMA is suitable for both confirmatory hypothesis tests and for exploratory analyses to ensure that no important effects were missed.
Both approaches can be included in a preregistration [@vanlissaComplementingPreregisteredConfirmatory2022].
With regard to data extraction,
it is important to strike a balance between inclusiveness and selectivity when coding moderators [@thompsonHowShouldMetaregression2002].
Moderators may include theoretically relevant factors and methodological ones, such as sample demographics, methods, instruments, study quality, and publication type.
A key challenge is that moderators may not always be reported.
The best way to handle missing data is by recovering the relevant information
by contacting authors or comparing different publications on the same data.
If data remains missing, users can use multiple imputation, which is a best practice for handling missingness (see Applied example).
Finally, effect sizes and their variances must be computed using suitable methods;
many of which are available in the R package `metafor` [@viechtbauerConductingMetaanalysesMetafor2010].

@@

With regard to publication, we highly recommend making the data and code for the meta-analysis publicly available.
One way to do this is by creating a reproducible research repository,
for example, using the Workflow for Reproducible Code in Science (WORCS) [@vanlissaWORCSWorkflowOpen2021].
Transparency allows readers and reviewers to verify that methods were correctly applied,
which bolsters confidence in the results.
Others can easily perform sensitivity analyses by changing the analysis code.
Sharing data allows the meta-analysis to be updated in the future,
which increases its reuse value.
Finally, sharing the model object (or code to reproduce it) allows others to obtain predictions for the expected effect size of a new study on the same topic.
These predictions can be used to conduct power analysis for future research.
To this end, researchers can simply enter their planned
design (or several alternative designs) as new lines of data, using the codebook of the
original meta-analysis, and use the published BRMA model to calculate the
predicted effect size for a study with these specifications.
